\textbf{Teaching a machine to generate an artificial universe}

### 2020-10-06 paper.tex
hi res kl graph, more sections

\textit{Architecture} --- 
Table~\ref{table:acgan_archi} of Appendix~\ref{appendix:acgan_setup} shows the architectures of the ACGAN discriminator 
and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
Appendix~\ref{appendix:acgan_setup}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. \\

{\setlength{\parindent}{0cm}
\textit{Training} --- 
The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Tensorflow~\cite{tensorflow} and 
Keras~\cite{keras}. The generator transforms a 256-dimensional vector drawn from a Gaussian prior $\mathbf{z}\sim 
\mathcal{N}(0,1)$ into a $32 \times 32 \times 32$ single-channel image. The latent-space vector is followed by a fully 
connected layer that is reshaped to a stack of feature maps for the convolutional layers. 
% Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} 
% for more detail on batch-normalisation and activation functions. 
}

@@

@@

\subsubsection{Universe generation}
% show three scales (side lengths in terms of generated cubes - e.g. 7, 9, 11) of generated-cube-universes.

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/mill_box_3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.19$.}
\label{fig:kernel_fig}
\end{figure}

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/gen_box_9_z3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.00$.}
\label{fig:kernel_fig}
\end{figure}

@@

\subsection{Loss and $D_{KL}$ curves}
% separation of D and G losses because of number of images shown to models in one training step

@@

@@

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_train.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples. This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all}
\end{figure*}

@@

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_unseen.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples with cube samples from unseen redshift snapshots in the Millennium-SAGE simulation 
         data This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all_z2}
\end{figure*}

@@

@@

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/graphs/z_kl.png}
\centering
\caption{The KL-trial means $\langle D_{KL} \rangle$ for comparisons of image distributions 
         $t$ and $g$ for all redshifts $z$ and $z'$ }
\label{fig:z_kl}
\end{figure}

@@

The condensed spectrum of representations for different features in the dataset of cosmic web at different redshift is, by
the ACGAN generator, mapped in an entangled way. The dimensions of $\mathbf{z}$ may not correspond to semantic features in
the data. There is no default ordering to the process through training. This means further work is required to find the 
effects of each possible axis in the latent space as well as the spherical interpolation paths in the latent space 
navigation.

@@

>>>>

\subsubsection{Architecture}
Table~\ref{table:acgan_archi} of Appendix~\ref{appendix:acgan_setup} shows the architectures of the ACGAN discriminator 
and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
Appendix~\ref{appendix:acgan_setup}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. 
% \textit{Architecture} --- 
% Table~\ref{table:acgan_archi} of Appendix~\ref{appendix:acgan_setup} shows the architectures of the ACGAN discriminator 
% and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
% Appendix~\ref{appendix:acgan_setup}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
% parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
% for the discriminator model. \\

% {\setlength{\parindent}{0cm}
% \textit{Training} --- 
% The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Tensorflow~\cite{tensorflow} and 
% Keras~\cite{keras}. The generator transforms a 256-dimensional vector drawn from a Gaussian prior $\mathbf{z}\sim 
% \mathcal{N}(0,1)$ into a $32 \times 32 \times 32$ single-channel image. The latent-space vector is followed by a fully 
% connected layer that is reshaped to a stack of feature maps for the convolutional layers. 
% % Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} 
% % for more detail on batch-normalisation and activation functions. 
% }

\subsubsection{Training}

The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Tensorflow~\cite{tensorflow} and 
Keras~\cite{keras}. The generator transforms a 256-dimensional vector drawn from a Gaussian prior $\mathbf{z}\sim 
\mathcal{N}(0,1)$ into a $32 \times 32 \times 32$ single-channel image. The latent-space vector is followed by a fully 
connected layer that is reshaped to a stack of feature maps for the convolutional layers. 
% Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} 
% for more detail on batch-normalisation and activation functions.

@@

\subsubsection{Spectral Normalization}
To control discriminator training and the heuristic feedback to the generator and its products, a method proposed by 
\cite{spectral_norm} uses the normalization of weight matrices in the discriminator and generator convolutional nets. This 
method allows control over the derivative of the discriminator value function in the adversarial setting of GANs. The 
Lipschitz constant of the discriminator is constrained without specifics concerning the inputs to either models. This 
data-independent regularization is an advantage over traditional regularization particularly in the generative context. 
The spectral norm of a weight matrix $W$ for a given layer $g : \mathbf{h}_{in} \rightarrow \mathbf{h}_{out}$ is

\begin{equation}
    \sigma(W) = \max_{\mathbf{h}:\mathbf{h}\neq 0} \frac{|| W \mathbf{h}||_2}{||\mathbf{h}||_2}
              = \max_{||\mathbf{h}||_2 \leq 1} ||W\mathbf{h}||_2
\end{equation}

which is equivalent to the singular value of $W$. This is applied to all layers in the discriminator and the generator.

@@

% \subsubsection{Universe generation}
% % show three scales (side lengths in terms of generated cubes - e.g. 7, 9, 11) of generated-cube-universes.

% \begin{figure}[!ht]%[hbt!]
% \includegraphics[width=\columnwidth]{figures/cubes/mill_box_3.png}
% \centering
% \caption{A box made of generated cubes at redshift $z=3.19$.}
% \label{fig:kernel_fig}
% \end{figure}

% \begin{figure}[!ht]%[hbt!]
% \includegraphics[width=\columnwidth]{figures/cubes/gen_box_9_z3.png}
% \centering
% \caption{A box made of generated cubes at redshift $z=3.00$.}
% \label{fig:kernel_fig}
% \end{figure}

@@

\subsection{Loss curves}
% separation of D and G losses because of number of images shown to models in one training step

@@

\subsection{$D_{KL}$ curves}

@@

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_train1.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples. This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all}
\end{figure*}

@@

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_unseen1.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples with cube samples from unseen redshift snapshots in the Millennium-SAGE simulation 
         data This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all_z2}
\end{figure*}

@@

The last two unseen redshift values show consistency between kernel sizes and they are similar values to the trained redshift 
values either side of these redshifts. Because these $\langle D_KL \rangle$ values have the lowest separation between the 
nearest trained redshifts, it suggests that the ACGAN has a certain range in redshift over which it can obtain the same coherence 
in samples with. The other redshifts suggest this boundary away from a given trained redshift value is between 0.3 and 0.5.

@@

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/graphs/z_kl1.png}
\centering
\caption{The KL-trial means $\langle D_{KL} \rangle$ for comparisons of image distributions 
         $t$ and $g$ for all redshifts $z$ and $z'$ }
\label{fig:z_kl}
\end{figure}

@@

The spectrum of representations for different features in the dataset is held in the latent space of the generator. The 
various features of the cosmic web at different redshift is mapped in an entangled way. The dimensions of $\mathbf{z}$ may 
not correspond to semantic features in the data. There is no default ordering to the process through training. This means 
further work is required to find the effects of each possible axis in the latent space as well as the spherical interpolation paths in the latent space navigation.

@@

More recent work~\cite{latent_disentanglement} provides another step towards solving the latent space entanglement problem.
Further work is required to explore the potential for these methods to this work, but there are no obvious reasons why this
would not fit into the current method.

\subsection{ACGAN architecture and training}
% SBN AND CBN
Batch normalisation~\cite{batchnorm} is critical to the function of the ACGAN presented in this work. However this method was derived for non-generative networks, and so normalising over a batch of conditioned latent vectors $\mathbf{z}(c)$ can distort the task of creating the latent space mapping for the generator. Conditional batch normalisation (CBN)~\cite{GAN_CBN} and self-modulated batch normalisation (SBN)~\cite{GAN_SBN} offer alternatives for this task by conditioning a set of learned parameters over the class space $C$ for normalisation. The difference between CBN and SBN are that the labels are required for the normalisation parameters in CBN whereas SBN depends only on the latent vectors. In practice CBN would be useful in both the discriminator and the generator.
% CGD
Recent work by Sch\"{a}fer et al.~\cite{GAN_CGD1, GAN_CGD2} provides an additional regularization, known as competitive gradient descent, to reinvent the interaction of the discriminator and generator models whilst maintaining their independent agencies. No metrics dependent on the space of training images are required such as in gradient penalty techniques for the same purpose. The method ensures that the discriminator model does not exploit the imperceptible errors of the generator products. This means that the generator loss is maximally related to the difference between the model and data distributions.
